---
title: "FARS3"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{model_details}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The functions programmed using data from the US National Highway Traffic Safety Administration's Fatality Analysis Reporting System (FARS), which is a nationwide census providing the American public yearly data regarding fatal injuries suffered in motor vehicle traffic crashes. 

### Libraries used in this project

```{r error=FALSE, warning=FALSE, message=FALSE, results='hide'}

library(dplyr)
library(readr)
library(magrittr)
library(tidyr)

```

### Function 1. This function reads data from .csv file

```{r error=FALSE, warning=FALSE, message=FALSE, results='hide'}

fars_read <- function(filename) {
  if(!file.exists(filename))
    stop("!file '", filename, "' does not exist")
  data <- suppressMessages({
    readr::read_csv(filename, progress = FALSE)
  })
  dplyr::tibble(data)
}

```

### Function's example:

```{r error=FALSE, warning=FALSE, message=FALSE, results='hide'}

filename <- "accident_2013.csv.bz2"
data <- fars_read(filename)

```

### Function 2. Make data file name

```{r error=FALSE, warning=FALSE, message=FALSE, results='hide'}

make_filename <- function(year) {
  year <- as.integer(year)
  sprintf("accident_%d.csv.bz2", year)
}

```

### Function's Example

```{r error=FALSE, warning=FALSE, message=FALSE, results='hide'}

make_filename(2015)

```

### Function 3. Read FARS years

```{r error=FALSE, warning=FALSE, message=FALSE, results='hide'}

fars_read_years <- function(years) {
  lapply(years, function(year) {
    file <- make_filename(year)
    tryCatch({
      dat <- fars_read(file)
      dat <- dat %>% mutate(year = YEAR) %>%
        select(MONTH, year) %>% data.frame()
    }, error = function(e) {
      warning("invalid year: ", year)
      return(NULL)
    })
  })
}

```

### Function's Example

```{r error=FALSE, warning=FALSE, message=FALSE, results='hide'}

fars_read_years(2013)

```

### Function 4. Summarize FARS data by years.Makes a summary count per month in particular years.This function summarizes yearly accidents data, by month.

```{r error=FALSE, warning=FALSE, message=FALSE, results='hide'}

fars_summarize_years <- function(years) {
  dat_list <- fars_read_years(years) %>% bind_rows()
  dat_list <- dat_list %>% group_by(year, MONTH) %>%
    summarise(n = n())
}

```

### Function's Examples

```{r error=FALSE, warning=FALSE, message=FALSE, results='hide'}

fars_summarize_years(2015)
fars_summarize_years(c(2015, 2014))

```

### Function 5. Display accidents map by state and year. Displays a plot with a state map including the accidents location by year.

```{r error=FALSE, warning=FALSE, message=FALSE, results='hide'}

fars_map_state <- function(state.num, year) {
  filename <- make_filename(year)
  data <- fars_read(filename)
  state.num <- as.integer(state.num)

  if(!(state.num %in% unique(data$STATE)))
    stop("invalid STATE number: ", state.num)
  data.sub <- dplyr::filter(data, STATE == state.num)
  if(nrow(data.sub) == 0L) {
    message("no accidents to plot")
    return(invisible(NULL))
  }
  is.na(data.sub$LONGITUD) <- data.sub$LONGITUD > 900
  is.na(data.sub$LATITUDE) <- data.sub$LATITUDE > 90
  with(data.sub, {
    maps::map("state", ylim = range(LATITUDE, na.rm = TRUE),
              xlim = range(LONGITUD, na.rm = TRUE))
    graphics::points(LONGITUD, LATITUDE, pch = 46)
  })
}

```

### Function's Example. It takes state #49 regarding 2015.

```{r error=FALSE, warning=FALSE, message=FALSE}

fars_map_state(49, 2015)

```
